---
title: "RL model results"
author: "Natalia Vélez, Yuan Chang Leong, Jamil Zaki, Hyowon Gweon"
date: "11/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(ggthemes)
require(RColorBrewer)
require(R.matlab)
dims = c('valence', 'setting', 'genre')
```

Data files:
```{r}
model_dir = 'model_RL'

# Self phase
self_dir = file.path(model_dir, 'self_phase')
self_files = list.files(self_dir)

# Training phase
train_dir = file.path(model_dir, 'training_phase_RL')
train_files = list.files(train_dir)

# Test phase
test_dir = file.path(model_dir, 'testing_phase_RL')
test_files = list.files(test_dir)

# Simulated agent weights
sim_dir = '../../inlab/sim_data/'
sim_files = list.files(sim_dir)
```

Load training data:
```{r}
train_data = NULL

for (f in train_files) {
  tmp_id = str_extract(f, 'sll_opusfmri_[0-9]+')
  tmp_run = str_extract(f, '(?<=train\\.)[0-9]')
  tmp_run = as.numeric(tmp_run)
  
  tmp_f = file.path(train_dir, f)
  
  tmp_data = read_tsv(tmp_f) %>%
    mutate(subject = tmp_id,
           run = tmp_run,
           trial = 1:n(),
           value_unchosen = ifelse(choice == 1, Value1, Value2),
           trial_type = 'train') %>%
    rename(value_chosen = Value_c,
           value_left = Value1,
           value_right = Value2,
           valence = W1,
           setting = W2,
           genre = W3,
           response_time = rt) %>%
    select(c(subject, run, trial, onset, duration, trial_type, response_time, choice, outcome, value_chosen, value_unchosen, everything()))
  
  train_data = plyr::rbind.fill(train_data, tmp_data)
  
   rm(list = grep('^tmp', ls(), value=T))
}
rm(list = c('f'))
```

Load test data:
```{r}
test_data = NULL

for (f in test_files) {
  tmp_id = str_extract(f, 'sll_opusfmri_[0-9]+')
  tmp_run = str_extract(f, '(?<=test\\.)[0-9]')
  tmp_run = as.numeric(tmp_run)
  
  tmp_f = file.path(train_dir, f)
  
  tmp_data = read_tsv(tmp_f) %>%
    mutate(subject = tmp_id,
           run = tmp_run,
           trial = 1:n(),
           value_unchosen = ifelse(choice == 1, Value1, Value2),
           trial_type = 'test') %>%
    rename(value_chosen = Value_c,
           value_left = Value1,
           value_right = Value2,
           valence = W1,
           setting = W2,
           genre = W3,
           response_time = rt) %>%
    select(c(subject, run, trial, onset, duration, trial_type, response_time, choice, outcome, value_chosen, value_unchosen, everything()))
  
  train_data = plyr::rbind.fill(train_data, tmp_data)
  
   rm(list = grep('^tmp', ls(), value=T))
}
rm(list = c('f'))
```

Load simulated agent weights:
```{r}
sim_data = NULL

for (f in sim_files) {
  tmp_f = file.path(sim_dir, f)
  tmp_id = str_extract(f, 'sll_opusfmri_[0-9]+')
  tmp_sim = readMat(tmp_f)$sim.agent
  tmp_weights = setNames(tmp_sim[,,1]$sim.weights, dims)
  
  tmp_data = data.frame(as.list(tmp_weights)) %>%
    mutate(subject = tmp_id) %>%
    select(c(subject, everything())) %>%
    gather(dimension, truth, valence:genre) %>%
    mutate(dimension = factor(dimension, levels = dims))
  
  sim_data = plyr::rbind.fill(sim_data, tmp_data)
  
  rm(list = grep('^tmp', ls(), value=T))
}

rm(list = c('f'))
```

Features only:
```{r}
train_features = train_data %>%
  group_by(subject) %>%
  mutate(trial = 1:n()) %>%
  ungroup() %>%
  select(c(subject, trial, valence:genre)) %>%
  gather(dimension, estimate, valence:genre) %>%
  mutate(dimension = factor(dimension, levels = dims)) %>%
  arrange(subject, trial, dimension)

head(train_features)
```

## Convergence during training phase
Accuracy during training phase:
```{r}
train_accuracy = train_data %>%
  drop_na(outcome) %>%
  group_by(subject) %>%
  summarise(accuracy = mean(outcome))
```

Plot weights over time:
```{r}
plot_convergence = function(s) {
  subj_train = train_features %>% filter(subject == s) %>% drop_na()
  subj_truth = sim_data %>% filter(subject == s)

  convergence_plot = ggplot(subj_train, aes(x = trial, y = estimate, color = dimension, group = dimension)) +
    geom_line(size=1.5) +
    geom_hline(data = subj_truth, aes(yintercept=truth, color = dimension), size=1, linetype = 'dashed') +
    theme_few(base_size = 18) +
    xlab('Trial') +
    ylab('Estimate') +
    ggtitle(sprintf('Subject: %s', s)) +
    scale_color_brewer('Dimension', labels = c('Valence', 'Setting', 'Genre'), palette = 'Set2') +
    theme(plot.title = element_text(hjust=0.5))
  
  return(convergence_plot)
}

subjects = sprintf('sll_opusfmri_%02d', 1:8)
for (s in subjects) {
  p = plot_convergence(s)
  ggsave(sprintf('plots/train_weights_%s.png', s), plot = p, width=6, height=4)
}

plot_convergence(subjects[1])
```

Cosine similarity between final model weights and true weights:
```{r}
cosine_similarity = function(v1, v2) {
  dotprod = sum(v1*v2)
  mag = sqrt(sum(v1^2)+sum(v2^2))
  sim = dotprod/mag
  return(sim)
}

train_final = train_features %>%
  group_by(subject) %>%
  filter(trial == max(trial)) %>%
  select(-trial) %>%
  left_join(sim_data, by=c('subject', 'dimension')) %>%
  summarise(similarity = cosine_similarity(estimate, truth)) %>%
  left_join(train_accuracy, by = 'subject')

train_final
```

Relate similarity to true weights vs. behavioral accuracy:
```{r}
ggplot(train_final, aes(x = accuracy, y = similarity)) +
  geom_point() +
  geom_smooth(method='lm', se=F) +
  theme_few(base_size=18) +
  xlab('Accuracy in training phase') +
  ylab('Similarity')
```

**Takeaway:** The more accurately participants respond during the training phase, the closer their responses are to the ground-truth values. (Also: everyone had above-chance performance in the training phase, which is nice!)